---
title: "GSB_script"
author: "Gabriel Silvestre Rocha"
date: "2023-02-26"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(ggplot2)
library(ggpubr)
library(knitr)
library(devtools)
library(ggbiplot)
library(corrplot)
library(readxl)
library(ARTool)
library(lsmeans)
library(ggeasy)
library(vegan)
library(factoextra)
library(FactoMineR)
library(multcomp)
library(tinytex)
```

# Data Upload
```{r}
setwd("C:/Users/gaaah/OneDrive/Área de Trabalho/Mestrado/Urochloa/resultados")

#Load main data
db <- readxl::read_xlsx("dbgeral2.xlsx", col_names = TRUE, col_types = NULL, na = "")

#Replace values
db$Depth[db$Depth == "40"] <- "30"

#Convert variables to factor
db$Depth <- as.factor(db$Depth)
db$Genotype <- as.factor(db$Genotype)

#Remove rows
dbc <- db[-c(49:60),]

#Load correlation data
dbcor <- readxl::read_xlsx("dbp.xlsx", col_names = TRUE, col_types = NULL, na = "")

#Convert variables to factor
dbcor$Depth <- as.factor(dbcor$Depth)
dbcor$Genotype <- as.factor(dbcor$Genotype)

#Remove rows
dbcor10c <- dbcor20c <- dbcor30c <- NULL

```

## Process Data for Depth = 10
```{r}
#Subset main data
res1 <- db[(db$Depth == "10"),]
db10 <- as.data.frame(res1)
db10$Depth = NULL
db10$Sample_ID = NULL
db10$Genotype <- as.factor(db10$Genotype)

#Remove rows
db10c <- db10[-c(17:20),]

#Subset correlation data
res1 <- dbcor[(dbcor$Depth == "10"),]
dbcor10 <- as.data.frame(res1)
dbcor10$Depth = NULL
dbcor10$Sample_ID = NULL
dbcor10$Genotype <- as.factor(dbcor10$Genotype)

#Remove rows
dbcor10c <- dbcor10[-c(17:20),]

```

## Process Data for Depth = 20
```{r}
#Subset main data
res2 <- db[(db$Depth == "20"),]
db20 <- as.data.frame(res2)
db20$Depth = NULL
db20$Sample_ID = NULL
db20$Genotype <- as.factor(db20$Genotype)

#Remove rows
db20c <- db20[-c(17:20),]

#Subset correlation data
res2 <- dbcor[(dbcor$Depth == "20"),]
dbcor20 <- as.data.frame(res2)
dbcor20$Depth = NULL
dbcor20$Sample_ID = NULL
dbcor20$Genotype <- as.factor(dbcor20$Genotype)

#Remove rows
dbcor20c <- dbcor20[-c(17:20),]

```

## Process Data for Depth = 30
```{r}
#Subset main data
res3 <- db[(db$Depth == "30"),]
db30 <- as.data.frame(res3)
db30$Depth = NULL
db30$Sample_ID = NULL
db30$Genotype <- as.factor(db30$Genotype)

#Remove rows
db30c <- db30[-c(17:20),]

#Subset correlation data
res3=(dbcor[(dbcor$Depth=="40"),])
dbcor30 <- as.data.frame(res3)
dbcor30$Depth = NULL
dbcor30$Sample_ID = NULL
dbcor30$Genotype <- as.factor(dbcor30$Genotype)

#Remove rows
dbcor30c = dbcor30[-c(17:20),]

```

# Ploting Graphs

## qPCR genes quantification

### Bacteria 16S
```{r}
bbac <- db %>%
  #filter(!(Genotype %in% c("Ipypora", "Control"))) %>%
  ggplot(aes(x=Genotype, y=bac2, fill=Genotype)) +
  geom_boxplot() +
  facet_grid(.~Depth, scales="free", space="free_x") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position="right",plot.title = element_text(size=11)) +
  ggtitle("Bacteria abundance") +
  xlab("Genotypes") +
  ylab("Bacteria 16S gene (gene copies/ng gDNA)")+
  ggeasy::easy_center_title() +
  ggeasy::easy_adjust_legend("center")
bbac
```

### Archaea 16S
```{r}
barc <- db %>%
  ggplot(aes(x=Genotype, y=archea, fill=Genotype)) +
  geom_boxplot() +
  facet_grid(.~Depth, scales="free", space="free_x") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position="right",plot.title = element_text(size=11)) +
  ggtitle("Archaea abundance") +
  xlab("Genotypes") +
  ylab("Archaea 16S gene (gene copies/ng gDNA)")+
  ggeasy::easy_center_title() +
  ggeasy::easy_adjust_legend("center")
barc
```

### Fungi ITS
```{r}
bits <- db %>%
  ggplot(aes(x=Genotype, y=ITS, fill=Genotype)) +
  geom_boxplot() +
  facet_grid(.~Depth, scales="free", space="free_x") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position="right",plot.title = element_text(size=11)) +
  ggtitle("Fungi abundance") +
  xlab("Genotypes") +
  ylab("Fungi ITS gene (gene copies/ng gDNA)")+
  ggeasy::easy_center_title() +
  ggeasy::easy_adjust_legend("center")
bits
```

### Alkaline phosphatase production by phoD gene
```{r}
bphod <- db %>%
  ggplot(aes(x=Genotype, y=phoD, fill=Genotype)) +
  geom_boxplot() +
  facet_grid(.~Depth, scales="free", space="free_x") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position="right",plot.title = element_text(size=11)) +
  ggtitle("phoD gene abundance") +
  xlab("Genotypes") +
  ylab("phoD gene (gene copies/ng gDNA)")+
  scale_y_log10()+
  ggeasy::easy_center_title() +
  ggeasy::easy_adjust_legend("center")
bphod
```

### P solubilization via pqqC
```{r}
bpqqc <- db %>%
  ggplot(aes(x=Genotype, y=pqqc3, fill=Genotype)) +
  geom_boxplot() +
  facet_grid(.~Depth, scales="free", space="free_x") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position="right",plot.title = element_text(size=11)) +
  ggtitle("pqqC gene abundance") +
  xlab("Genotypes") +
  ylab("pqqC gene (gene copies/ng gDNA)") +
  ggeasy::easy_center_title() +
  ggeasy::easy_adjust_legend("center")
bpqqc
```

### AMF colonization
```{r}
bfma <- db10c %>%
  ggplot(aes(x=Genotype, y=colonization, fill=Genotype)) +
  geom_boxplot() +
  scale_fill_manual(values=c("#4477AA", "#AA7744", "#E7B800", "#117733"))+
  geom_jitter(color="black", size=0.4, alpha=0.9) +
  theme_bw() +
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank()) +
  theme(legend.position="right",plot.title = element_text(size=11)) +
  ggtitle("AMF colonization") +
  xlab("Genotypes") +
  ylab("AMF colonization (%)")+
  ggeasy::easy_center_title() +
  ggeasy::easy_adjust_legend("center")
bfma
```

## Plotting boxplots all together
```{r}
ggarrange(barc, bbac, bits, bphod, bpqqc, bfma,
          labels = c("A", "B", "C", "D", "E", "F"),
          common.legend = TRUE,
          legend = "bottom",
          ncol = 3, nrow = 2)
```

## Correlation plots
### pqqC gene
```{r}
pixpqq<-ggplot(db, aes(pqqc3, NaOHi)) + 
  geom_point(size = 4) +
  #facet_wrap(db$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 1200, label.y = 80) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 1200, label.y = 75, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  ylab("Inorganic P (mg/Kg)") +
  xlab("pqqC gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
pixpqq

pixpqq2<-ggplot(db, aes(pqqc3, NaOHi)) + 
  geom_point(size = 4) +
  facet_wrap(db$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 10, label.y = 85) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 10, label.y = 78, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  ylab("Inorganic P (mg/Kg)") +
  xlab("pqqC gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
pixpqq2

arcxpqq<-ggplot(db, aes(pqqc2, archea)) + 
  geom_point(size = 4) +
  #facet_wrap(db$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 50, label.y = 32000) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 50, label.y = 30000, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  ylab("Archaea 16S gene (gene copies/ng gDNA)") +
  xlab("pqqC gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
arcxpqq

arcxpqq2<-ggplot(db, aes(pqqc2, archea)) + 
  geom_point(size = 4) +
  facet_wrap(db$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 50, label.y = 32000) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 50, label.y = 30000, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  ylab("Archaea 16S gene (gene copies/ng gDNA)") +
  xlab("pqqC gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
arcxpqq2
```

## Plotting all together
```{r}
ggarrange(pixpqq,pixpqq2, arcxpqq, arcxpqq2,
          labels = c("A", "B", "C","D"),
          common.legend = TRUE,
          legend = "none",
          ncol = 2, nrow = 2)
```

### phoD gene
```{r}
pdxfal<-ggplot(db, aes(fosfalk, phoD)) + 
  geom_point(size = 4) +
  #facet_wrap(dbc$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 20, label.y = 78000) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 20, label.y = 72000, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  xlab("Alk. phosp. acitivty (ug p-nitrofenol/g h-1)") +
  ylab("phoD gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
pdxfal

pdxfal2<-ggplot(db, aes(fosfalk, phoD)) + 
  geom_point(size = 4) +
  facet_wrap(db$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 20, label.y = 78000) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 20, label.y = 72000, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  xlab("Alk. phosp. acitivty (ug p-nitrofenol/g h-1)") +
  ylab("phoD gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
pdxfal2

pdxpi<-ggplot(db, aes(NaOHi, phoD)) + 
  geom_point(size = 4) +
  #facet_wrap(db$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 10, label.y = 78000) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 10, label.y = 72000, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  xlab("Inorganic P (mg/kg)") +
  ylab("phoD gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
pdxpi

pdxpi2<-ggplot(db, aes(NaOHi, phoD)) + 
  geom_point(size = 4) +
  facet_wrap(db$Genotype, nrow = 1)+
  geom_smooth(method = lm, se =T, color = "black") +
  stat_cor(aes(label = paste(..r.label.., ..p.label.., sep = "~`,`~")),
           r.accuracy = 0.01,
           p.accuracy = 0.01, size = 4,
           label.x = 10, label.y = 78000) +
  stat_regline_equation(aes(label = ..eq.label..),
                        label.x = 10, label.y = 72000, size = 4) +
  theme_bw()+
  theme(legend.position="right",plot.title = element_text(size=11), axis.text.y=element_text(size = 13), axis.text.x=element_text(size = 13)) +
  ggtitle("") +
  xlab("Inorganic P (mg/kg)") +
  ylab("phoD gene (gene copies/ng gDNA)") +
  scale_fill_manual(values=c("#8E6698", "#4477AA", "#AA7744", "#E7B800", "#117733"))
pdxpi2
```

## Plotting all together
```{r}
ggarrange(pdxfal, pdxfal2, pdxpi, pdxpi2,
          labels = c("A", "B", "C", "D"),
          common.legend = TRUE,
          legend = "none",
          ncol = 2, nrow = 2)
```

# Statistical analysis
## AMF colonization
```{r}
stpqq <- lm(colonization ~ Genotype, data = db10c)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)
```

## Bacteria 16S
```{r}
#10cm
stpqq <- lm(bac2 ~ Genotype, data = db10)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)

#30cm
stpqq <- lm(bac2 ~ Genotype, data = db30)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)
```

## Archaea 16S
```{r}
#10cm
stpqq <- lm(archea ~ Genotype, data = db10)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)

#20cm
stpqq <- lm(archea ~ Genotype, data = db20)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)

#30cm
stpqq <- lm(archea ~ Genotype, data = db30)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)
```

## Fungi ITS
```{r}
#20cm
stpqq <- lm(ITS ~ Genotype, data = db20)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)

#30cm
stpqq <- lm(ITS ~ Genotype, data = db30)
mc <- glht(stpqq, linfct = mcp(Genotype = "Tukey"))
mc_letters <- cld(mc, Letters = letters)
print(mc_letters)
```

# Principal Components Analysis (PCA)
### Depth: 0-10cm
```{r}
pca1 <- PCA(scale(dbcor10[,c(4,8,10:11,13,15:17,19:22)]), scale = TRUE)
fviz_pca_biplot(pca1, habillage = dbcor10$Genotype, 
                addEllipses = TRUE, 
                pointshape = 19,
                pointsize = 2,
                geom = c("point"),
                ggtheme = theme_pubr(),
                palette = c("#8E6698","#4477AA", "#AA7744", "#E7B800", "#117733"),
                ellipse.type = "convex", col.var = "black")

factors1 <- as.data.frame(as.factor(dbcor10$Genotype))
dbcor10_2 <- as.data.frame(dbcor10[,c(4,8,10:11,13,15:17,19:22)])
vegan::adonis2(dbcor10_2 ~ factors1[c(1:20),], permutations = 999, method = "bray")

# Contributions of variables to PC1
fviz_contrib(pca1, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(pca1, choice = "var", axes = 2, top = 5)
```

### Depth: 10-20cm
```{r}
pca20 <- PCA(scale(dbcor20[,c(4,8,10:11,13,15:17,19:22)]), scale = TRUE)
fviz_pca_biplot(pca20, habillage = dbcor10$Genotype, 
                addEllipses = TRUE, 
                pointshape = 19,
                pointsize = 2,
                geom = c("point"),
                ggtheme = theme_pubr(),
                palette = c("#8E6698","#4477AA", "#AA7744", "#E7B800", "#117733"),
                ellipse.type = "convex", col.var = "black")

factors20 <- as.data.frame(as.factor(dbcor20$Genotype))
dbcor20_2 <- as.data.frame(dbcor20[,c(4,8,10:11,13,15:17,19:22)])
vegan::adonis2(dbcor20_2 ~ factors20[c(1:20),], permutations = 999, method = "bray") #p = 0.001 ***

# Contributions of variables to PC1
fviz_contrib(pca20, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(pca20, choice = "var", axes = 2, top = 5)

```

### Depth: 20-30cm
```{r}
pca30 <- PCA(scale(dbcor30[,c(4,8,10:11,13,15:17,19:22)]), scale = TRUE)
fviz_pca_biplot(pca30, habillage = dbcor10$Genotype, 
                addEllipses = TRUE, 
                pointshape = 19,
                pointsize = 2,
                geom = c("point"),
                ggtheme = theme_pubr(),
                palette = c("#8E6698","#4477AA", "#AA7744", "#E7B800", "#117733"),
                ellipse.type = "convex", col.var = "black")

factors30 <- as.data.frame(as.factor(dbcor30$Genotype))
dbcor30_2 <- as.data.frame(dbcor30[,c(4,8,10:11,13,15:17,19:22)])
vegan::adonis2(dbcor30_2 ~ factors30[c(1:20),], permutations = 999, method = "bray") #p = 0.001 ***

# Contributions of variables to PC1
fviz_contrib(pca30, choice = "var", axes = 1, top = 5)
# Contributions of variables to PC2
fviz_contrib(pca30, choice = "var", axes = 2, top = 5)
```

